---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading data, echo=TRUE}

data<-read.csv("activity.csv",header=TRUE)
head(data)

```
```{r preprocessing data, echo=TRUE}


datanna <- data[!is.na(data$steps),]

```
## What is mean total number of steps taken per day?
### 1. Make a histogram of the total number of steps taken each day

```{r mean total, fig.width=8, fig.height=6, echo=TRUE}
sumStepsByDay <- aggregate(datanna$steps, by=list(date = datanna$date), FUN=sum)
names(sumStepsByDay)[2]<-"steps"
hist(sumStepsByDay$steps,main="Total number of steps taken each day",xlab="Total number of steps")
```

### 2.Calculate and report the mean and median total number of steps taken per day
```{r report, echo = TRUE}
meanSteps<-aggregate(datanna$steps, by=list(date = datanna$date), FUN=function(x) c(mean=mean(x),median=median(x)))

meanSteps

```

## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r average by time series , echo= TRUE, fig.height=6, fig.width= 10 }
aveStepByTime <- aggregate(datanna$steps, by=list(interval = datanna$interval), FUN=mean)
names(aveStepByTime)[2] <- "ave.steps"
 with(aveStepByTime,
      plot(interval,ave.steps,
           main="Average Daily Activity", 
           xlab="time interval", 
           ylab="Average steps",
           type="l")
 )
 
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxnum number}
aveStepByTime[order(aveStepByTime$ave.steps,decreasing=T)[1],] 
 
```

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
``` {r number of NAs, echo=TRUE }
sum(is.na(data))
 
```

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

``` {r filling missing value using mean for that 5-minute interval , echo = TRUE}

 newData <- data
 
 for ( row in 1:nrow(data)){
  if( is.na(newData[row,'steps']) )
    newData[row,'steps'] <- 
      aveStepByTime[aveStepByTime$interval == data[row,'interval'],'ave.steps']
  
}
 
```

### 4. Make a histogram of the total number of steps taken each day 
```{r sum total, fig.width=8, fig.height=6, echo=TRUE}
newSumStepsByDay <- aggregate(newData$steps, by=list(date = newData$date), FUN=sum)
names(newSumStepsByDay)[2]<-"steps"
hist(newSumStepsByDay$steps,main="Total number of steps taken each day",xlab="Total number of steps")
```
### Calculate and report the mean and median total number of steps taken per day
```{r new report, echo = TRUE}
newMeanSteps<-aggregate(newData$steps, by=list(date = newData$date), FUN=function(x) c(mean=mean(x),median=median(x)))

newMeanSteps

```
### Do these values differ from the estimates from the first part of the assignment
```{r differ report, echo = TRUE}
differ <- merge(newMeanSteps,meanSteps, by= "date" , all.x=TRUE)
differ

```

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r weekdays, echo = TRUE}
Sys.setlocale("LC_TIME", "English") 
newData['weekdays'] <- sapply(as.Date(newData$date),
                              function(x) if(weekdays(x)== "Sunday" |
                                             weekdays(x)=="Saturday") 0 
                                          else 1 )

```

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r weekdays plot, echo = TRUE, fig.width=10, fig.height=6 }
weekdaysMeanByTime <- aggregate(newData$steps,
                                by=list(interval = newData$interval, 
                                        weekdays = newData$weekdays),
                                FUN=mean)

library(lattice)

with ( weekdaysMeanByTime,
      xyplot(main="Activity patterns",
       x~interval|factor(weekdays,levels=c(0,1),
  	   labels=c("weekend","weekday")) ,
       ylab ="Number of steps",
       xlab = "Interval",
       type = "l",
       layout=c(1,2)
      )
)

```


